---
title: "WTMS Nicely Formatted tables"
output: 
  html_notebook:
    toc: true

---



# Load the Libraries
```{r}
# Need this package for DPLYR and ggplot etc.
install.packages("tidyverse")
library(tidyverse)

# Need this package for formating tables.
install.packages("formattable")
library(formattable)

```

# Load the Data

```{r}
Issues_2020.08.11 <- read.csv("Issues_2020-08-11.csv")
Furniture_2020.08.11 <- read.csv("Furniture_2020-08-11.csv")
```


I started this Notebook as it became obvious that data structure needs to be different depending on whether I wanted to output Tables or Charts.

# Data Transformation

Here we need to get the data into a single table for display. I'm going to try and start simple with the "Present Furniture" of the RMP Tables. This is a summary table of WCP furniture (not including Signs).

We'll start off by defining the Table categories for Present Furniture:

## The Template
```{r}
RMP_PresentFurniture <- data.frame(RMP_Group = c("Boardwalks", "Bridges", "Gates Kissing", "Gates Other", "Steps - Flights", "Stiles - Ladder", "Stiles - Step", "Stiles - Stone"))
```

## Issues

### Select relevant variables and create category variables
First we need to tidy up the table by selecting only the variables we're interested in:

```{r}
Issues <- select(Issues_2020.08.11, Region, IssueType, IssueTypeGroup, StatusDesc, PriorityDesc, IssueCode, ManagingAuth, Complete, Firm, ProtrusionFree, SafelySited, HandpostPresent, StepsInPlace, SwingFreely, LatchesAndBolts, HorseLatch, BackpackNegotiable, LogoPresent, OrientedCorrectly, ClearlyVisible, Bilingual, RouteName, Action, OverallDateLastUpdated, PromotedRoute)
```


Then we need to do the hard job of assigning the Furniture and Issue records to each of the above categories using the Furniture or Issue Types:

```{r}
# The below is NOT PERFECT, and I'm starting to think it is better to define these categories in CAMS - this would allow greated understanding of which furniture items are scored within RMPs for local authorities if/when they get access to CAMS Web.
Issues <- mutate(Issues,
       RMP_Group = case_when(
  grepl("Boardwalk", IssueType) ~ "Boardwalks",
  grepl("Bridge", IssueType) ~ "Bridges",
  grepl("Gate", IssueType) ~ "Gates Other",
  grepl("Kissing", IssueType) ~ "Gates Kissing",
  grepl("Steps", IssueType) ~ "Steps - Flights",
  grepl("Ladder", IssueType) ~ "Stiles - Ladder",
  grepl("Step", IssueType) ~ "Stiles - Step",
  grepl("Stone", IssueType) ~ "Stiles - Stone"
  )) %>% 

# This is a good way of identifying categories which are reported under QS, we can then filter
mutate(Issues,
       QS_Tables = if_else(grepl("[*]$",IssueType), "Yes", "No"))

```

We also need to tidy up the Quality Standards variables.

```{r}
# Let's see if we can use mutate to just replace "Pass" "Fail" and NA with numeric values.

Issues <- mutate(Issues, 
                 Complete = if_else(grepl("Fail",Complete), "Fail", "Pass"),
                 Firm = if_else(grepl("Fail",Firm), "Fail", "Pass"),
                 ProtrusionFree = if_else(grepl("Fail",ProtrusionFree), "Fail", "Pass"),
                 SafelySited = if_else(grepl("Fail",SafelySited), "Fail", "Pass"),
                 HandpostPresent = if_else(grepl("Fail",HandpostPresent), "Fail", "Pass"),
                 StepsInPlace = if_else(grepl("Fail",StepsInPlace), "Fail", "Pass"),
                 SwingFreely = if_else(grepl("Fail",SwingFreely), "Fail", "Pass"),
                 LatchesAndBolts = if_else(grepl("Fail",LatchesAndBolts), "Fail", "Pass"),
                 BackpackNegotiable = if_else(grepl("Fail",BackpackNegotiable), "Fail", "Pass"),
                 LogoPresent = if_else(grepl("Fail",LogoPresent), "Fail", "Pass"),
                 OrientedCorrectly = if_else(grepl("Fail",OrientedCorrectly), "Fail", "Pass"),
                 ClearlyVisible = if_else(grepl("Fail",ClearlyVisible), "Fail", "Pass"),
                 Bilingual = if_else(grepl("Fail",Bilingual), "Fail", "Pass"),
                 RouteName = if_else(grepl("Fail",RouteName), "Fail", "Pass"),
                 )

```

Then we need to filter out only the relevant records

```{r}
# Now lets filter only for the Issue records we're interested in reporting.
Issues_FurniturePresent <-   filter(Issues, QS_Tables == "Yes" & StatusDesc == "Unresolved" & !is.na(RMP_Group))
```


### Issues Summary

### Number of Issues

Now we're gonna create a Summary Table for Issues 



```{r}
Issues_Summary_Table <- 
  Issues_FurniturePresent %>% 
  group_by(RMP_Group) %>% 
  summarise(Issues = n())

Issues_Summary_Table
```

### Count Quality Standard test Fails

```{r}
QS_Summary_Table <- data.frame(RMP_Group = Issues_FurniturePresent)
```

filter(Issues_FurniturePresent, Complete == "Fail") %>% 
```{r}
Complete <-  Issues_FurniturePresent %>% 
  filter(Complete == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(Complete = n())

Firm <-  Issues_FurniturePresent %>% 
  filter(Firm == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(Firm = n())

ProtrusionFree <-  Issues_FurniturePresent %>% 
  filter(ProtrusionFree == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(ProtrusionFree = n())

SafelySited <-  Issues_FurniturePresent %>% 
  filter(SafelySited == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(SafelySited = n())

HandpostPresent <-  Issues_FurniturePresent %>% 
  filter(HandpostPresent == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(HandpostPresent = n())

StepsInPlace <-  Issues_FurniturePresent %>% 
  filter(StepsInPlace == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(StepsInPlace = n())

SwingFreely <-  Issues_FurniturePresent %>% 
  filter(SwingFreely == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(SwingFreely = n())

LatchesAndBolts <-  Issues_FurniturePresent %>% 
  filter(LatchesAndBolts == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(LatchesAndBolts = n())

HorseLatch <-  Issues_FurniturePresent %>% 
  filter(HorseLatch == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(HorseLatch = n())

BackpackNegotiable <-  Issues_FurniturePresent %>% 
  filter(BackpackNegotiable == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(BackpackNegotiable = n())

LogoPresent <-  Issues_FurniturePresent %>% 
  filter(LogoPresent == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(LogoPresent = n())

OrientedCorrectly <-  Issues_FurniturePresent %>% 
  filter(OrientedCorrectly == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(OrientedCorrectly = n())

ClearlyVisible <-  Issues_FurniturePresent %>% 
  filter(ClearlyVisible == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(ClearlyVisible = n())

Bilingual <-  Issues_FurniturePresent %>% 
  filter(Bilingual == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(Bilingual = n())

RouteName <-  Issues_FurniturePresent %>% 
  filter(RouteName == "Fail") %>% 
  group_by(RMP_Group) %>%
  summarise(RouteName = n())

QS_Summary_Table <- left_join(RMP_PresentFurniture, Complete, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, Firm, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, ProtrusionFree, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, SafelySited, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, HandpostPresent, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, StepsInPlace, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, SwingFreely, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, LatchesAndBolts, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, HorseLatch, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, BackpackNegotiable, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, LogoPresent, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, OrientedCorrectly, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, ClearlyVisible, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, Bilingual, by = "RMP_Group")
QS_Summary_Table <-   left_join(QS_Summary_Table, RouteName, by = "RMP_Group")

QS_Summary_Table

rm(Complete, Firm, ProtrusionFree, SafelySited, HandpostPresent, StepsInPlace, SwingFreely, LatchesAndBolts, HorseLatch, BackpackNegotiable, LogoPresent, OrientedCorrectly, ClearlyVisible, Bilingual, RouteName)
```


## Furniture

As before, we first need to create a tidy data frame by selecting only the variables we're interested in:

```{r}
Furniture <- select(Furniture_2020.08.11, Region, FurnitureType, FurnitureGroupType, LatestSurveyCondition, InfraCode)
```

Then, as with Issues, we need to assign RMP categories.

```{r}
# See previous comment under "Issues" for various reasons I think this would be better defined with CAMS
Furniture <- mutate(Furniture,
       RMP_Group = case_when(
  grepl("Boardwalk", FurnitureType) ~ "Boardwalks",
  grepl("Bridge", FurnitureType) ~ "Bridges",
  grepl("Gate", FurnitureType) ~ "Gates Other",
  grepl("Kissing", FurnitureType) ~ "Gates Kissing",
  grepl("Steps", FurnitureType) ~ "Steps - Flights",
  grepl("Ladder", FurnitureType) ~ "Stiles - Ladder",
  grepl("Step", FurnitureType) ~ "Stiles - Step",
  grepl("Stone", FurnitureType) ~ "Stiles - Stone"
  )) %>% 

# This is a good way of identifying categories which are reported under QS, we can then filter
mutate(Furniture,
       QS_Tables = if_else(grepl("[*]$",FurnitureType), "Yes", "No"))

# Now lets filter only for the Issue records we're interested in reporting.
Furniture_FurniturePresent <-   filter(Furniture, QS_Tables == "Yes" & LatestSurveyCondition != "Not Present" & !is.na(RMP_Group))


```


Now we're gonna create a Summary Table for Furniture

```{r}
Furniture_Summary_Table <- 
  Furniture_FurniturePresent %>% 
  group_by(RMP_Group) %>% 
  summarise(Furniture = n())

Furniture_Summary_Table
```

## Merging Furniture and Issue summaries

Now for the challenging bit!

```{r}
# Let's initially try to use merge function (which can only merge two data frames at a time).

Furniture_Present_Summary <- left_join(RMP_PresentFurniture, Furniture_Summary_Table, by = "RMP_Group")

Furniture_Present_Summary <-   left_join(Furniture_Present_Summary, Issues_Summary_Table, by = "RMP_Group")

Furniture_Present_Summary
```

# Final Table Options

## Overall Number of Furniture Vs Number of Issues

```{r}
Furniture_Present_Summary <- mutate(Furniture_Present_Summary, IssueVsFurn = Issues / Furniture)

Furniture_Present_Summary[is.na(Furniture_Present_Summary)] <- 0

FPS_Formattable <- data.frame(
  RMP_Group = Furniture_Present_Summary$RMP_Group,
  Furniture = Furniture_Present_Summary$Furniture,
  Issues = Furniture_Present_Summary$Issues,
  IssueVsFurn = percent(Furniture_Present_Summary$IssueVsFurn)
)


formattable(FPS_Formattable,
            align = c("l", "c", "c", "r"), 
            list(
              RMP_Group = formatter("span", style = ~ style(color = "grey",font.weight = "bold", width = "200")), 
              IssueVsFurn = color_bar("lightblue"),
              Furniture = color_tile("transparent", "Grey"),
              Issues = color_tile("transparent", "pink")
                 ))


```

## Quality Standards Tests

```{r}
# Formattable doesn't like NA's so this switches them for 0s
QS_Summary_Table[is.na(QS_Summary_Table)] <- 0
```

### Boardwalks

```{r}
#Boardwalks
QS_Summary_Table_Boardwalks <- select(QS_Summary_Table, RMP_Group, Complete, Firm, ProtrusionFree, SafelySited)
QS_Summary_Table_Boardwalks <- filter(QS_Summary_Table_Boardwalks, RMP_Group %in% c("Boardwalks"))

formattable(QS_Summary_Table_Boardwalks, 
            align = c("l", rep("c", NCOL(QS_Summary_Table_Boardwalks))),
            list(
  RMP_Group = formatter("span", style = ~ style(color = "grey",font.weight = "bold")),              
  Complete = color_tile("white", "pink"),
  Firm = color_tile("white", "pink"),
  ProtrusionFree = color_tile("white", "pink"),
  SafelySited = color_tile("white", "pink")
            ))
rm(QS_Summary_Table_Boardwalks)
```


### Bridges

```{r}
QS_Summary_Table_Bridges <- select(QS_Summary_Table, RMP_Group, Complete, Firm, ProtrusionFree, SafelySited, LogoPresent)
QS_Summary_Table_Bridges <- filter(QS_Summary_Table_Bridges, RMP_Group %in% c("Bridges"))

formattable(QS_Summary_Table_Bridges, 
            align = c("l", rep("c", NCOL(QS_Summary_Table_Bridges))),
            list(
  RMP_Group = formatter("span", style = ~ style(color = "grey",font.weight = "bold")),              
  Complete = color_tile("white", "pink"),
  Firm = color_tile("white", "pink"),
  ProtrusionFree = color_tile("white", "pink"),
  SafelySited = color_tile("white", "pink"),
  LogoPresent = color_tile("white", "pink")
            ))
rm(QS_Summary_Table_Bridges)
```


### Gates

```{r}
# Gates
QS_Summary_Table_Gates <- select(QS_Summary_Table, RMP_Group, Complete, Firm, ProtrusionFree, SafelySited, SwingFreely, LatchesAndBolts, HorseLatch, BackpackNegotiable, LogoPresent)
QS_Summary_Table_Gates <- filter(QS_Summary_Table_Gates, RMP_Group %in% c("Gates Kissing", "Gates Other"))

formattable(QS_Summary_Table_Gates, 
            align = c("l", rep("c", NCOL(QS_Summary_Table_Gates))),
            list(
  RMP_Group = formatter("span", style = ~ style(color = "grey",font.weight = "bold")),              
  Complete = color_tile("white", "pink"),
  Firm = color_tile("white", "pink"),
  ProtrusionFree = color_tile("white", "pink"),
  SafelySited = color_tile("white", "pink"),
  SwingFreely = color_tile("white", "pink"),
  LatchesAndBolts = color_tile("white", "pink"),
  HorseLatch = color_tile("white", "pink"),
  BackpackNegotiable = color_tile("white", "pink"),
  LogoPresent = color_tile("white", "pink")
            ))
rm(QS_Summary_Table_Gates)
            
```


### Stiles

```{r}
# Stiles
QS_Summary_Table_Stiles <- select(QS_Summary_Table, RMP_Group, Complete, Firm, ProtrusionFree, SafelySited, HandpostPresent, StepsInPlace, LogoPresent)
QS_Summary_Table_Stiles <- filter(QS_Summary_Table_Stiles, RMP_Group %in% c("Stiles - Ladder", "Stiles - Step", "Stiles - Stone"))


formattable(QS_Summary_Table_Stiles, 
            align = c("l", rep("c", NCOL(QS_Summary_Table_Stiles))),
            list(
  RMP_Group = formatter("span", style = ~ style(color = "grey",font.weight = "bold")),              
  Complete = color_tile("white", "pink"),
  Firm = color_tile("white", "pink"),
  ProtrusionFree = color_tile("white", "pink"),
  SafelySited = color_tile("white", "pink"),
  HandpostPresent = color_tile("white", "pink"),
  StepsInPlace = color_tile("white", "pink"),
  LogoPresent = color_tile("white", "pink")
    ))

rm(QS_Summary_Table_Stiles)
```


```{r}







# Signs
formattable(QS_Summary_Table, 
            align = c("l", rep("c", NCOL(QS_Summary_Table))),
            list(
  RMP_Group = formatter("span", style = ~ style(color = "grey",font.weight = "bold")),              
  Complete = color_tile("white", "pink"),
  Firm = color_tile("white", "pink"),
  ProtrusionFree = color_tile("white", "pink"),
  SafelySited = color_tile("white", "pink"),
OrientedCorrectly = color_tile("white", "pink"),
  ClearlyVisible = color_tile("white", "pink"),
  Bilingual = color_tile("white", "pink"),
  RouteName = color_tile("white", "pink")

))

# 
```


# Themes

## Formattable

Initially here I'd like to explore the potential for using Formattable: (https://renkun-ken.github.io/formattable/)
Some good resources for Formattable here:
- https://cran.r-project.org/web/packages/formattable/vignettes/formattable-data-frame.html
- https://www.littlemissdata.com/blog/prettytables
- https://renkun-ken.github.io/formattable/
- https://www.displayr.com/formattable/
- https://bioinfo.iric.ca/create-a-nice-looking-table-using-r/



```{r}
install.packages("formattable")
library(formattable)
```

Here's a nice example script for Formattable from the above website:



## DataTables

Might be worth checking out DataTables at some point, if I want to create interactive web tables

- https://rstudio.github.io/DT/

## Shiny

A teeny bit optimistic, but could give shiny a go!
